---
title: "Functions"
author: "Erica Goto"
date: "3/11/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Functions

```{r dataset}
#library(xlsx)
library(dplyr)
library(readxl)
library(ggplot2)
escorregamento <- read_excel("BD-IPT 2010_todos.xls")

df1 <- read_excel("IPT_SP_2010.xls")
df2 <- read_excel("sector_slope_angle.xlsx")

merge_df <- merge(df1, df2, by.x = "setor", 
             by.y = "sg_setor_r", all.x = TRUE, all.y = TRUE)

#write.xlsx(merge_df, "merge_df.xls")

```


### Excluding  cells with empty values
1 Find natural slope angle using 5m DEM
2) Raster to Point so I could do Spatial Join to the polygons of the risk sectors
3) New polygon of risk sectors and natural slope angle to excel table (so I could merge the sectors with the AHP calculation)
4) excel table: need to exclude NA crated with the merge of the two excel file
5) created just one column with slope angle (both anthropogenic and natural). The ones that did not have a value for the anthropogenic I inputed the natural one. So mainly the angle that AHP will be using is the anthropogenic and if there were not change into the terrain, the natural one. 
```{r}
escorregamento <- read_excel("merge_df_les.xls")

```






#Changing values data,frane

```{r , include=FALSE}

# these values were extracted from classifiers.xls These weight were normalized with z-score


escorregamento$brick <- ifelse(escorregamento$brick == 1, -0.847144412, escorregamento$brick)
escorregamento$wood <- ifelse(escorregamento$wood == 1, 0.018621387, escorregamento$wood)
escorregamento$mixed<- ifelse(escorregamento$mixed == 1, 0.180267533, escorregamento$mixed)


escorregamento$d1 <- ifelse(escorregamento$d1 == 1, -0.847144412, escorregamento$d1)
escorregamento$d2 <- ifelse(escorregamento$d2 == 1, -0.173939960802013, escorregamento$d2)
escorregamento$d3 <- ifelse(escorregamento$d3 == 1, -0.00104045691529944, escorregamento$d3)
escorregamento$d4<- ifelse(escorregamento$d4 == 1, -0.291658771958924, escorregamento$d4)

escorregamento$garbage<- ifelse(escorregamento$garbage == 1, 0.260848862548364, escorregamento$garbage)
escorregamento$land_fill<- ifelse(escorregamento$land_fill == 1, 0.260848862548364, escorregamento$land_fill)
escorregamento$construction_material<- ifelse(escorregamento$construction_material == 1, 0.260848862548364, escorregamento$construction_material)
escorregamento$garbage_all <- (escorregamento$garbage+escorregamento$land_fill+escorregamento$construction_material)
escorregamento$garbage_all <- ifelse(escorregamento$garbage_all >= 0.260848862548364, 0.260848862548364, 0)


escorregamento$cracks <- ifelse(escorregamento$cracks == 1, 1.31713866640692, escorregamento$cracks) # crack in the house
escorregamento$fracture <- ifelse(escorregamento$fracture == 1, 2.34774965617479, escorregamento$fracture) # crack in the teraain
escorregamento$tilted <- ifelse(escorregamento$tilted == 1, 1.4717303148721, escorregamento$tilted)
escorregamento$landslide_scars <- ifelse(escorregamento$landslide_scars == 1, 2.22751170736854, escorregamento$landslide_scars)
escorregamento$belly_wall <- ifelse(escorregamento$belly_wall == 1, 1.4717303148721, escorregamento$belly_wall)
escorregamento$downward_floor  <- ifelse(escorregamento$downward_floor == 1, 3.25812269713641, escorregamento$downward_floor)

escorregamento$drainage_none <- ifelse(escorregamento$drainage_none == 1, -0.668824645993864, escorregamento$drainage_none)
escorregamento$drainage_precarious <- ifelse(escorregamento$drainage_precarious == 1, -0.699186045020083, escorregamento$drainage_precarious)
escorregamento$drainage_okay <- ifelse(escorregamento$drainage_okay == 1, -0.847144412105597, escorregamento$drainage_okay)


escorregamento$tree <- ifelse(escorregamento$tree == 1, -0.809970465644889, escorregamento$tree)
escorregamento$shrubs <- ifelse(escorregamento$shrubs == 1, -0.83785092549042, escorregamento$shrubs)
escorregamento$banana_tree <- ifelse(escorregamento$banana_tree == 1, 0.0186213873767418, escorregamento$banana_tree)
escorregamento$desnude_soil <- ifelse(escorregamento$desnude_soil == 1, 0.114731452565228, escorregamento$desnude_soil)



escorregamento$septic_tank <- ifelse(escorregamento$septic_tank == 1, -0.607280655919903, escorregamento$septic_tank)
escorregamento$leaking <- ifelse(escorregamento$leaking == 1, -0.290092399384874, escorregamento$leaking)
escorregamento$wastewater <- ifelse(escorregamento$wastewater == 1, -0.572563533314079, escorregamento$wastewater)
escorregamento$rainfall_water_conc <- ifelse(escorregamento$rainfall_water_conc == 1, -0.567829380231467, escorregamento$rainfall_water_conc)

escorregamento$angle_both <- as.numeric(escorregamento$angle_both)





#this need to be in a decreasing order otherwise one will change the other
escorregamento$angle_both  <- ifelse(escorregamento$angle_both <(10), 0, escorregamento$angle_both)
escorregamento$angle_both  <- ifelse(escorregamento$angle_both >=(10) & escorregamento$angle_both < 17, 0.026693875, escorregamento$angle_both)
escorregamento$angle_both  <- ifelse(escorregamento$angle_both >=(17) & escorregamento$angle_both < 30, 0.013041282, escorregamento$angle_both)
escorregamento$angle_both  <- ifelse(escorregamento$angle_both >=(30) & escorregamento$angle_both < 60, 0.026693875, escorregamento$angle_both)
escorregamento$angle_both  <- ifelse(escorregamento$angle_both  >= (60) & escorregamento$angle_both  < 90, 0.05664807, escorregamento$angle_both)
escorregamento$angle_both  <- ifelse(escorregamento$angle_both  == 90, 0.058482, escorregamento$angle_both)


escorregamento$natural_slope  <- ifelse(escorregamento$natural_slope >=(10) & escorregamento$natural_slope < 17, 0.026693875, escorregamento$natural_slope)
escorregamento$natural_slope  <- ifelse(escorregamento$natural_slope >=(17) & escorregamento$natural_slope < 30, 0.013041282, escorregamento$natural_slope)
escorregamento$natural_slope  <- ifelse(escorregamento$natural_slope >=(30) & escorregamento$natural_slope < 60, 0.026693875, escorregamento$natural_slope)
escorregamento$natural_slope  <- ifelse(escorregamento$natural_slope  >= (60) & escorregamento$natural_slope  < 90, 0.05664807, escorregamento$natural_slope)
escorregamento$natural_slope  <- ifelse(escorregamento$natural_slope  == 90, 0.058482, escorregamento$natural_slope)




escorregamento<- escorregamento %>%
  mutate(value=(brick+wood+mixed+d1+d2+d3+d4+cracks+fracture+downward_floor+tilted+landslide_scars+belly_wall+ tree + shrubs + banana_tree + desnude_soil + septic_tank+ leaking + wastewater+ rainfall_water_conc + garbage_all + drainage_okay + drainage_precarious + drainage_none) + angle_both) %>%
   mutate(value_natural_angle =(brick+wood+mixed+d1+d2+d3+d4+cracks+fracture+downward_floor+tilted+landslide_scars+belly_wall+ tree + shrubs + banana_tree + desnude_soil + septic_tank+ leaking + wastewater+ rainfall_water_conc + garbage_all + drainage_okay + drainage_precarious + drainage_none) + natural_slope)


    
```

## Ploting risk value
```{r}

sd_value <- sd(escorregamento$value, na.rm=TRUE)
mean_value <- mean(escorregamento$value, na.rm=TRUE)

ggplot(escorregamento, aes(value)) + 
  geom_histogram(aes(y=..density..), colour="black", fill="steelblue4",binwidth=0.2) +
  geom_density(alpha=0.2, fill="gray")+
      ggtitle("Risk Value")  +
  geom_vline(aes(xintercept=mean(value, na.rm=TRUE)+0.5*sd(value, na.rm=TRUE)),
            color="firebrick4", linetype="dashed", size=1) +
   geom_vline(aes(xintercept=mean(value, na.rm=TRUE)- 0.5*sd(value, na.rm=TRUE)),
            color="firebrick4", linetype="dashed", size=1) +
   geom_vline(aes(xintercept=mean(value, na.rm=TRUE)+1.5*sd(value, na.rm=TRUE)),
            color="gray46", linetype="dashed", size=1) +
   geom_vline(aes(xintercept=mean(value, na.rm=TRUE)-1.5*sd(value, na.rm=TRUE)),
            color="gray46", linetype="dashed", size=1)

## natural slope

sd_value <- sd(escorregamento$value_natural_angle, na.rm=TRUE)
mean_value <- mean(escorregamento$value_natural_angle, na.rm=TRUE)

ggplot(escorregamento, aes(value_natural_angle)) + 
  geom_histogram(aes(y=..density..), colour="black", fill="steelblue4",binwidth=0.2) +
  geom_density(alpha=0.2, fill="gray")+
      ggtitle("Risk Value")  +
  geom_vline(aes(xintercept=mean(value_natural_angle, na.rm=TRUE)),
            color="firebrick4", linetype="dashed", size=1) +
   geom_vline(aes(xintercept=mean(value_natural_angle, na.rm=TRUE)+sd(value_natural_angle, na.rm=TRUE)),
            color="gray46", linetype="dashed", size=1) +
   geom_vline(aes(xintercept=mean(value_natural_angle, na.rm=TRUE)-sd(value_natural_angle, na.rm=TRUE)),
            color="gray46", linetype="dashed", size=1)


```



## Counting number of values to decide threshold
```{r}
### 
sd_value <- sd(escorregamento$value, na.rm=TRUE)
mean_value <- mean(escorregamento$value, na.rm=TRUE)

perc <- escorregamento$value[!is.na(escorregamento$value)]
 
A <- length(perc)

B <- sum(perc< (mean(perc)-sd(perc, na.rm=TRUE)))
C <- sum(perc< mean(perc) & perc >= mean(perc)+(-sd(perc)))
D <- sum (perc  >= mean(perc) & perc< sd(perc)+mean(perc))
E <- sum(perc >= sd(perc)+mean(perc))

B+C+D+E 

df_risk_level_perc <- data.frame(
  c("low", "moderate", "high", "very high"),
  c((B/A),(C/A), (D/A), (E/A) )
)

df_risk_level_perc


### for rik level with natural angle

sd_value <- sd(escorregamento$value_natural_angle, na.rm=TRUE)
mean_value <- mean(escorregamento$value_natural_angle, na.rm=TRUE)

perc <- escorregamento$value_natural_angle[!is.na(escorregamento$value_natural_angle)]
 
A <- length(perc)

B <- sum(perc< (mean(perc)-sd(perc, na.rm=TRUE)))
C <- sum(perc< mean(perc) & perc >= mean(perc)+(-sd(perc)))
D <- sum (perc  >= mean(perc) & perc< sd(perc)+mean(perc))
E <- sum(perc >= sd(perc)+mean(perc))

B+C+D+E 

df_risk_level_perc <- data.frame(
  c("low", "moderate", "high", "very high"),
  c((B/A),(C/A), (D/A), (E/A) )
)

df_risk_level_perc

```

## changing the values for risk level 
```{r}
## Categorical level for vulnerability

new_escorregamento <- escorregamento %>%
  select(sector, value, value_natural_angle) %>%
  mutate(risk_level = value)

new_escorregamento <- na.omit(new_escorregamento) # exclude rows with NA


### using cut() to create the intervals


new_escorregamento$risk_level_one_sd <- cut(new_escorregamento$risk_level, breaks=c(-6,(mean_value-sd_value), mean_value, (mean_value + sd_value),7), labels = c("low", "moderate", "high", "very high" ))
new_escorregamento$risk_level_half_sd <- cut(new_escorregamento$risk_level, breaks=c(-6, (mean_value-0.5*sd_value), (mean_value + 0.5*sd_value), (mean_value + 1.5*sd_value), 7), labels = c("low", "moderate", "high", "very high" ))

new_escorregamento$risk_level_natural_slope_one_sd <- cut(new_escorregamento$risk_level, breaks=c(-6,(mean_value-sd_value), mean_value, (mean_value + sd_value),7), labels = c("low", "moderate", "high", "very high" ))
new_escorregamento$risk_level_natural_slope_half_sd <- cut(new_escorregamento$risk_level, breaks=c(-6, (mean_value-0.5*sd_value), (mean_value + 0.5*sd_value), (mean_value + 1.5*sd_value), 16), labels = c("low", "moderate", "high", "very high" ))


library ("openxlsx")

#write.xlsx(new_escorregamento, file="risk_value_AHP.xls")

```

### Open merged file



### Merging risk_level(shapefile) with risk_value_AHP
```{r}
library(rgdal)

# The input file geodatabase


fgdb <- "shapefile/social_Vulnerability_sp.gdb"

# List all feature classes in a file geodatabase
subset(ogrDrivers(), grepl("GDB", name))
fc_list <- ogrListLayers(fgdb)
print(fc_list)



vuln <- readOGR(dsn=fgdb,layer="to_be_merged_with_risk_level")



vuln@data <- merge(vuln@data,new_escorregamento, by.x="sg_setor_r", by.y="sector", all.x=T, sort=F)

names(vuln@data)


## create the laye
#writeOGR(census_tract_sp,dsn="shapefile",layer="census_tract_sp_vuln_correct2", driver="ESRI Shapefile")


```