---
title: "Functions"
author: "Erica Goto"
date: "3/11/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## In this file I computed the weight again. Did some correction and adjustments. 
1) removed soil, geology, and land cover since they do not appear in the dataset I am using to compute the weight
2) value of position of building is added to all risk sectors with one value. 
3) description of the weight can be found in the weight_normalized.xls tab: removed one not used

```{r dataset}
#library(xlsx)
library(dplyr)
library(readxl)
library(ggplot2)
escorregamento <- read_excel("BD-IPT 2010_todos.xls")

df1 <- read_excel("IPT_SP_2010.xls")
df2 <- read_excel("sector_slope_angle.xlsx")

merge_df <- merge(df1, df2, by.x = "setor", 
             by.y = "sg_setor_r", all.x = TRUE, all.y = TRUE)

#write.xlsx(merge_df, "merge_df.xls")

```


### Excluding  cells with empty values
1 Find natural slope angle using 5m DEM
2) Raster to Point so I could do Spatial Join to the polygons of the risk sectors
3) New polygon of risk sectors and natural slope angle to excel table (so I could merge the sectors with the AHP calculation)
4) excel table: need to exclude NA crated with the merge of the two excel file
5) created just one column with slope angle (both anthropogenic and natural). The ones that did not have a value for the anthropogenic I inputed the natural one. So mainly the angle that AHP will be using is the anthropogenic and if there were not change into the terrain, the natural one. 
```{r}
escorregamento <- read_excel("merge_df_les.xls")

```






#Changing values data,frane

```{r , include=FALSE}

# these values were extracted from weight_normalized.xls tab: remove ones not used

#Type of building (wood = 0.030040224, brikc=0, mixed-material:0.035649)
escorregamento$brick <- ifelse(escorregamento$brick == 1, 0, escorregamento$brick)
escorregamento$wood <- ifelse(escorregamento$wood == 1, 0.030040224, escorregamento$wood)
escorregamento$mixed<- ifelse(escorregamento$mixed == 1, 0.035649, escorregamento$mixed)

#Density (d1=0, d2=0.023358757, d3=0.029358, d4=0.019274165)
escorregamento$d1 <- ifelse(escorregamento$d1 == 1, 0, escorregamento$d1)
escorregamento$d2 <- ifelse(escorregamento$d2 == 1, 0.023358757, escorregamento$d2)
escorregamento$d3 <- ifelse(escorregamento$d3 == 1, 0.029358, escorregamento$d3)
escorregamento$d4<- ifelse(escorregamento$d4 == 1, 0.019274165, escorregamento$d4)

# garbage=0.038445
escorregamento$garbage<- ifelse(escorregamento$garbage == 1, 0.038445, escorregamento$garbage)
escorregamento$land_fill<- ifelse(escorregamento$land_fill == 1, 0.038445, escorregamento$land_fill)
escorregamento$construction_material<- ifelse(escorregamento$construction_material == 1, 0.038445, escorregamento$construction_material)


escorregamento$garbage_all <- (escorregamento$garbage+escorregamento$land_fill+escorregamento$construction_material)
escorregamento$garbage_all <- ifelse(escorregamento$garbage_all >= 0.038445, 0.038445, 0)

#instability (leaning_wall=0.08046, crack_house=0.075096, tree_pole_tilted = 0.08046, downward_floor = 0.142444, landslide_scar = 0.106684, crak_terrain= 0.110856)

escorregamento$cracks <- ifelse(escorregamento$cracks == 1, 0.075096, escorregamento$cracks) # crack in the house
escorregamento$fracture <- ifelse(escorregamento$fracture == 1, 0.110856, escorregamento$fracture) # crack in the teraain
escorregamento$tilted <- ifelse(escorregamento$tilted == 1, 0.08046, escorregamento$tilted)
escorregamento$landslide_scars <- ifelse(escorregamento$landslide_scars == 1, 0.106684, escorregamento$landslide_scars)
escorregamento$belly_wall <- ifelse(escorregamento$belly_wall == 1, 0.08046, escorregamento$belly_wall)
escorregamento$downward_floor  <- ifelse(escorregamento$downward_floor == 1, 0.142444, escorregamento$downward_floor)

#drainage (none=0.006187315, precarious =0.005133839, okay=0)

escorregamento$drainage_none <- ifelse(escorregamento$drainage_none == 1, 0.006187315, escorregamento$drainage_none)
escorregamento$drainage_precarious <- ifelse(escorregamento$drainage_precarious == 1, 0.005133839, escorregamento$drainage_precarious)
escorregamento$drainage_okay <- ifelse(escorregamento$drainage_okay == 1, 0, escorregamento$drainage_okay)

#land use (tree= 0.003668352, shrub = 0.00343908, soil=0.01127254, grass=0.00401226, banban_tree=0.0124189, urbanized=0.003362656)
escorregamento$tree <- ifelse(escorregamento$tree == 1, 0.003668352, escorregamento$tree)
escorregamento$shrubs <- ifelse(escorregamento$shrubs == 1, 0.00343908, escorregamento$shrubs)
escorregamento$banana_tree <- ifelse(escorregamento$banana_tree == 1, 0.0124189, escorregamento$banana_tree)
escorregamento$desnude_soil <- ifelse(escorregamento$desnude_soil == 1, 0.01127254, escorregamento$desnude_soil)

#add grass to all sectors since it does not have the option = 0.00401226


#Drainage(wastewater= 0.00952737, rain_water = 0.009691635, leaking = 0.019328515, septic_tank= 0.00996541)

escorregamento$septic_tank <- ifelse(escorregamento$septic_tank == 1, 0.00996541, escorregamento$septic_tank)
escorregamento$leaking <- ifelse(escorregamento$leaking == 1, 0.019328515, escorregamento$leaking)
escorregamento$wastewater <- ifelse(escorregamento$wastewater == 1, 0.00952737, escorregamento$wastewater)
escorregamento$rainfall_water_conc <- ifelse(escorregamento$rainfall_water_conc == 1, 0.009691635, escorregamento$rainfall_water_conc)

escorregamento$angle_both <- as.numeric(escorregamento$angle_both)

escorregamento$natural_slope <- as.numeric(escorregamento$natural_slope)

escorregamento$natural_slope  <- ifelse(escorregamento$natural_slope <(10), 0, escorregamento$natural_slope)
escorregamento$natural_slope  <- ifelse(escorregamento$natural_slope >=(10) & escorregamento$natural_slope < 17, 0.011916376, escorregamento$natural_slope)
escorregamento$natural_slope  <- ifelse(escorregamento$natural_slope >=(17) & escorregamento$natural_slope < 30, 0.038132404, escorregamento$natural_slope)
escorregamento$natural_slope  <- ifelse(escorregamento$natural_slope >=(30) & escorregamento$natural_slope < 60, 0.078052265, escorregamento$natural_slope)
escorregamento$natural_slope  <- ifelse(escorregamento$natural_slope  >= (60) & escorregamento$natural_slope  < 90, 0.165637631, escorregamento$natural_slope)
escorregamento$natural_slope  <- ifelse(escorregamento$natural_slope  == 90, 0.171, escorregamento$natural_slope)





#this need to be in a decreasing order otherwise one will change the other
escorregamento$angle_both  <- ifelse(escorregamento$angle_both <(10), 0, escorregamento$angle_both)
escorregamento$angle_both  <- ifelse(escorregamento$angle_both >=(10) & escorregamento$angle_both < 17, 0.011916376, escorregamento$angle_both)
escorregamento$angle_both  <- ifelse(escorregamento$angle_both >=(17) & escorregamento$angle_both < 30, 0.038132404, escorregamento$angle_both)
escorregamento$angle_both  <- ifelse(escorregamento$angle_both >=(30) & escorregamento$angle_both < 60, 0.078052265, escorregamento$angle_both)
escorregamento$angle_both  <- ifelse(escorregamento$angle_both  >= (60) & escorregamento$angle_both  < 90, 0.165637631, escorregamento$angle_both)
escorregamento$angle_both  <- ifelse(escorregamento$angle_both  == 90, 0.171, escorregamento$angle_both)



#both grass and building location value added to all

escorregamento<- escorregamento %>%
  mutate(value=(brick+wood+mixed+d1+d2+d3+d4+cracks+fracture+downward_floor+tilted+landslide_scars+belly_wall+ tree + shrubs + banana_tree + desnude_soil + septic_tank+ leaking + wastewater+ rainfall_water_conc + garbage_all + drainage_okay + drainage_precarious + drainage_none) + angle_both+0.00401226+0.036814  ) %>%
   mutate(value_natural_angle =(brick+wood+mixed+d1+d2+d3+d4+cracks+fracture+downward_floor+tilted+landslide_scars+belly_wall+ tree + shrubs + banana_tree + desnude_soil + septic_tank+ leaking + wastewater+ rainfall_water_conc + garbage_all + drainage_okay + drainage_precarious + drainage_none) + natural_slope + 0.00401226 + 0.036814 ) %>%
  mutate(value_instability=(cracks+fracture+downward_floor+tilted+landslide_scars+belly_wall))%>%
  mutate(value_anthropogenic=(brick+wood+mixed+d1+d2+d3+d4+tree + shrubs + banana_tree + desnude_soil+ septic_tank+ leaking + wastewater+ rainfall_water_conc + garbage_all + drainage_okay + drainage_precarious + drainage_none)) %>%
  mutate(value_natural=(natural_slope+0.00401226+ 0.036814 ))




    
```

## Ploting risk value
```{r}

sd_value <- sd(escorregamento$value, na.rm=TRUE)
mean_value <- mean(escorregamento$value, na.rm=TRUE)

ggplot(escorregamento, aes(value)) + 
  geom_histogram(aes(y=..density..), colour="black", fill="steelblue4",binwidth=0.2) +
  geom_density(alpha=0.2, fill="gray")+
      ggtitle("Risk Value")  +
  geom_vline(aes(xintercept=mean(value, na.rm=TRUE)+0.5*sd(value, na.rm=TRUE)),
            color="firebrick4", linetype="dashed", size=1) +
   geom_vline(aes(xintercept=mean(value, na.rm=TRUE)- 0.5*sd(value, na.rm=TRUE)),
            color="firebrick4", linetype="dashed", size=1) +
   geom_vline(aes(xintercept=mean(value, na.rm=TRUE)+1.5*sd(value, na.rm=TRUE)),
            color="gray46", linetype="dashed", size=1) +
   geom_vline(aes(xintercept=mean(value, na.rm=TRUE)-1.5*sd(value, na.rm=TRUE)),
            color="gray46", linetype="dashed", size=1)

## natural slope

sd_value_natural <- sd(escorregamento$value_natural_angle, na.rm=TRUE)
mean_value_natural <- mean(escorregamento$value_natural_angle, na.rm=TRUE)

ggplot(escorregamento, aes(value_natural_angle)) + 
  geom_histogram(aes(y=..density..), colour="black", fill="steelblue4",binwidth=0.2) +
  geom_density(alpha=0.2, fill="gray")+
      ggtitle("Risk Value")  +
  geom_vline(aes(xintercept=mean(value_natural_angle, na.rm=TRUE)),
            color="firebrick4", linetype="dashed", size=1) +
   geom_vline(aes(xintercept=mean(value_natural_angle, na.rm=TRUE)+sd(value_natural_angle, na.rm=TRUE)),
            color="gray46", linetype="dashed", size=1) +
   geom_vline(aes(xintercept=mean(value_natural_angle, na.rm=TRUE)-sd(value_natural_angle, na.rm=TRUE)),
            color="gray46", linetype="dashed", size=1)


```



## Counting number of values to decide threshold
```{r}
### 
sd_value <- sd(escorregamento$value, na.rm=TRUE)
mean_value <- mean(escorregamento$value, na.rm=TRUE)

perc <- escorregamento$value[!is.na(escorregamento$value)]
 
A <- length(perc)

B <- sum(perc< (mean(perc)-sd(perc, na.rm=TRUE)))
C <- sum(perc< mean(perc) & perc >= mean(perc)+(-sd(perc)))
D <- sum (perc  >= mean(perc) & perc< sd(perc)+mean(perc))
E <- sum(perc >= sd(perc)+mean(perc))

B+C+D+E 

df_risk_level_perc <- data.frame(
  c("low", "moderate", "high", "very high"),
  c((B/A),(C/A), (D/A), (E/A) )
)

df_risk_level_perc


### for rik level with natural angle

sd_value_natural <- sd(escorregamento$value_natural_angle, na.rm=TRUE)
mean_value_natural <- mean(escorregamento$value_natural_angle, na.rm=TRUE)

perc <- escorregamento$value_natural_angle[!is.na(escorregamento$value_natural_angle)]
 
A <- length(perc)

B <- sum(perc< (mean(perc)-sd(perc, na.rm=TRUE)))
C <- sum(perc< mean(perc) & perc >= mean(perc)+(-sd(perc)))
D <- sum (perc  >= mean(perc) & perc< sd(perc)+mean(perc))
E <- sum(perc >= sd(perc)+mean(perc))

B+C+D+E 

df_risk_level_perc <- data.frame(
  c("low", "moderate", "high", "very high"),
  c((B/A),(C/A), (D/A), (E/A) )
)

df_risk_level_perc

```

## changing the values for risk level 
```{r}
## Categorical level for vulnerability

new_escorregamento <- escorregamento %>%
  select(sector, value, value_natural_angle, value_anthropogenic, value_instability, value_natural) %>%
  mutate(risk_level = value)

new_escorregamento <- na.omit(new_escorregamento) # exclude rows with NA


### using cut() to create the intervals


new_escorregamento$risk_level_one_sd <- cut(new_escorregamento$risk_level, breaks=c(-6,(mean_value-sd_value), mean_value, (mean_value + sd_value),7), labels = c("low", "moderate", "high", "very high" ))
new_escorregamento$risk_level_half_sd <- cut(new_escorregamento$risk_level, breaks=c(-6, (mean_value-0.5*sd_value), (mean_value + 0.5*sd_value), (mean_value + 1.5*sd_value), 7), labels = c("low", "moderate", "high", "very high" ))

new_escorregamento$risk_level_natural_slope_one_sd <- cut(new_escorregamento$value_natural_angle, breaks=c(-6,(mean_value_natural-sd_value_natural), mean_value, (mean_value_natural + sd_value_natural),7), labels = c("low", "moderate", "high", "very high" ))
new_escorregamento$risk_level_natural_slope_half_sd <- cut(new_escorregamento$value_natural_angle, breaks=c(-6, (mean_value_natural-0.5*sd_value_natural), (mean_value_natural + 0.5*sd_value_natural), (mean_value_natural + 1.5*sd_value_natural), 16), labels = c("low", "moderate", "high", "very high" ))


library ("openxlsx")

#write.xlsx(new_escorregamento, file="risk_value_AHP_updated.xls")

```


## Creating 4 cagegories from 5 since BGM uses low, moderate, high, and very high. Therefore, low and very low are going to computed but aggreaged as low. 
- using cut as mean-sd, mean-0.5sd, mean+0.5sd, mean+sd
- with four categories: mean-0.5sd, mean+0.5sd, mean+sd

```{r}
### using cut() to create the intervals


new_escorregamento$new_risk_level_both_slopes <- cut(new_escorregamento$risk_level, breaks=c(-6,(mean_value-0.5*sd_value), (mean_value+0.5*sd_value), (mean_value + sd_value),7), labels = c("low", "moderate", "high", "very high" ))


new_escorregamento$new_risk_level_natural <- cut(new_escorregamento$value_natural_angle, breaks=c(-6, (mean_value_natural-0.5*sd_value_natural), (mean_value_natural + 0.5*sd_value_natural), (mean_value_natural +sd_value_natural), 7), labels = c("low", "moderate", "high", "very high" ))


```

###Importing dataset from arcgis to merge with the one created
```{r}
df <- read_excel("risk_level_updated_look_into_R.xlsx")

df1 <- df %>%
  select(nm_area_ri:cd_grau_ri)

new_df <- merge(df1, new_escorregamento, by.x = "sg_setor_r", 
             by.y = "sector", all.x = TRUE, all.y = FALSE)

library ("openxlsx")

#write.xlsx(new_df, file="risk_value_final_version20210330.xlsx")


```

