---
title: "Untitled"
author: "E G"
date: "October 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
###IPT Training Set Weight - Divided in 3 (AHP, Ordinary Logistic Regression, Validation)

```{r}

library(tidyverse)
library(dplyr)
library(readxl)
library(effsize)
IPT <- read_excel("ipt_test.xlsx")

IPT_todo <- read_excel("IPT_SP_2010.xls")
AHP_Training <- read_excel("ipt_training_set_AHP_BGM.xlsx")


```

Changing AHP Training to compute Risk Value
```{r}
AHP <- AHP_Training %>% 
  select(setor, Inclinacao, Total_Moradia, Densidade, GeolDesf, Solo, CoberturaNatural, AterroLixoEntulho, InstabSum, AguaTotal, VegTotal, Risco)

AHP$Total_Moradia <- as.numeric(AHP$Total_Moradia)
AHP$Densidade <- as.numeric(AHP$Densidade)
AHP$GeolDesf <- as.numeric(AHP$GeolDesf)
AHP$Solo <- as.numeric(AHP$Solo)
AHP$CoberturaNatural <- as.numeric(AHP$CoberturaNatural)
AHP$AterroLixoEntulho <- as.numeric(AHP$AterroLixoEntulho)
AHP$InstabSum <- as.numeric(AHP$InstabSum)
AHP$AguaTotal <- as.numeric(AHP$AguaTotal)
AHP$VegTotal <- as.numeric(AHP$VegTotal)
AHP$Inclinacao <- as.numeric(AHP$Inclinacao)


#change values InstabSum, AguaTotal, VegTogal


AHP$InstabSum <- ifelse(AHP$InstabSum > 23.7, 23.7, AHP$InstabSum)
AHP$AguaTotal <- ifelse(AHP$AguaTotal > 6.0424868, 6.0424868, AHP$AguaTotal)
AHP$VegTotal <- ifelse(AHP$VegTotal > 3.8590672, 3.8590672, AHP$VegTotal)

AHP1 <- AHP  %>% 
  mutate(RiskNum = Inclinacao + Total_Moradia + Densidade + GeolDesf + Solo + CoberturaNatural + AterroLixoEntulho +  InstabSum +  AguaTotal + VegTotal) %>% 
select(setor, Risco, RiskNum)

AHP1 <- AHP1 %>%
  mutate(LogRisk = log10(RiskNum))

colnames(AHP1) <- c("Sector", "Risk", "RiskNum", "LogRisk")

AHP1


```

Ploting AHP1 to look the distribution
```{r AHP1_plot}
Hist <- hist(AHP1$LogRisk, breaks = 80)
QQPlot <- qqnorm(AHP1$RiskNum)

par(mfrow=c(1,2))

BoxPlot <- ggplot(AHP1, aes(x=Risk, y=RiskNum)) +
  geom_boxplot(aes(fill=Risk)) +
  theme_bw()+
  ggtitle("Values for Each Degree of Risk")+
  xlab("Risk Degree")+
  ylab("Risk Value")+ 
    theme(plot.title = element_text(hjust = 0.5)) ##center
BoxPlot 

Jiterplot <- ggplot (AHP1, aes(x=Risk, y=RiskNum)) +
  geom_jitter(aes(colour = Risk), width = 0.3, alpha =0.4)+
  theme_classic ()+ 
      labs(x="Degree of Risk", y = "Risk Value") + 
    theme(plot.title = element_text(hjust = 0.5)) ##center
Jiterplot

hist_ggplot <- ggplot(AHP1, aes(x=RiskNum)) +
                        geom_histogram(aes(y=..density..), binwidth=1, color="dark green", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") + 
    labs(title="Risk value histogram",x="Risk Value", y = "Density") + 
    theme(plot.title = element_text(hjust = 0.5)) ##center


  
hist_ggplot + scale_x_continuous(limits = c(0, 80), breaks = c(0,5, 10, 15, 20, 25, 30, 35, 40, 45, 50, 55, 60, 65, 70, 75, 80))

# Values from the first normal distrubed from hist_ggplot (<53)

norm1 <- AHP1 %>% 
  filter(RiskNum <=53)

mean_norm1 <-mean(norm1$RiskNum)
mean_norm1 


norm2 <- AHP1 %>% 
  filter(RiskNum >53)

mean_norm2 <-mean(norm2$RiskNum)
mean_norm2 
  

```

Comparing Risk Degree to see if they are significant different

## Test 1 (R1 VS R2)
- H0: no significant difference between R1 and R2
- H1: There is a significant different between R1 and R2

## Test 2 (R2 VS R3)
- H0: no significant difference between R3 and R2
- H1: There is a significant different between R3 and R2

## Test 3 (R3 VS R4)
- H0: no significant difference between R3 and R4
- H1: There is a significant different between R3 and R4
```{r}

#Prepare dataset

R1<- AHP1%>%
  filter(Risk=="R1") %>%
  select(RiskNum)
R1<- R1$RiskNum

R2<- AHP1%>%
  filter(Risk=="R2") %>%
  select(RiskNum)
R2<- R2$RiskNum

R3<- AHP1%>%
  filter(Risk=="R3") %>%
  select(RiskNum)
R3<- R3$RiskNum

R4<- AHP1%>%
  filter(Risk=="R4") %>%
  select(RiskNum)
R4<- R4$RiskNum

t_test_diff_R1_R2 <- t.test (R1, R2, var.equal = TRUE)
t_test_diff_R1_R2
effect_size_R1_R2 <- cohen.d(R1, R2)
effect_size_R1_R2


t_test_diff_R2_R3 <- t.test (R2, R3, var.equal = TRUE)
t_test_diff_R2_R3
effect_size_R2_R3 <- cohen.d(R2, R3)
effect_size_R2_R3

t_test_diff_R3_R4 <- t.test (R3, R4, var.equal = TRUE)
t_test_diff_R3_R4
effect_size_R3_R4 <- cohen.d(R3, R4)
effect_size_R3_R4

```



Values Average & SD


```{r}
AHP_R1 <- AHP1 %>% 
  filter(Risk == "R1") 
AHP_R1_Mean <- mean(AHP_R1$RiskNum)
AHP_R1_SD <- sd(AHP_R1$RiskNum)

AHP_R2 <- AHP1 %>% 
  filter(Risk == "R2") 
AHP_R2_Mean <- mean(AHP_R2$RiskNum)
AHP_R2_SD <- sd(AHP_R2$RiskNum)

AHP_R3 <- AHP1 %>% 
  filter(Risk == "R3") 
AHP_R3_Mean <- mean(AHP_R3$RiskNum)
AHP_R3_SD <- sd(AHP_R3$RiskNum)

AHP_R4 <- AHP1 %>% 
  filter(Risk == "R4") 
AHP_R4_Mean <- mean(AHP_R4$RiskNum)
AHP_R4_SD <- sd(AHP_R4$RiskNum)

df_AHP <- data.frame(c( "R1", "R2", "R3", "R4"),
                  c(AHP_R1_Mean, AHP_R2_Mean, AHP_R3_Mean, AHP_R4_Mean), 
                 c( AHP_R1_SD, AHP_R2_SD, AHP_R3_SD, AHP_R4_SD))
colnames(df_AHP) <-c("Risk", "MEan", "SD")
df_AHP

plot(df_AHP$MEan)


```