---
title: "AHP Brazilian Government Methodology"
author: "Erica Goto"
date: "12/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library (dplyr)
```

## AHP Brazilian Goverment Methodology
### Part 01 - Anthrop, Natural or Instability

Dataset: Include row with the values

```{r dataset}

# data = row with the values for the matrix
# n= number of parameters
# MM = main matrix 
# MP = priority matrix (weight)
# Sum = sum of each column of the MM
# MN = normalized matrix
# AX = MP*MM
# Lamb_Max
# CI = critical index
# RI = random index
# CR = consistency ratio

# return: CR, MP


#'  AHP
#' 
#' Function computes weight and consistancy ratio
#' @param data vector to build the matrix
#' @param n number of criteria
#' @param RI random index
#' @param CI critical index
#' @param Lamb_Max
#' @return CR, MP Critical Ratio and Priority matrix (weight)

AHP <- function (data, n, MM, MN, MP, AX, Lamb_Max, CI, CR, RI) {

MM <- matrix(data, nrow=n, ncol=n, byrow = TRUE)

#Sum <- colSums (MM)
MN <- sweep(MM, FUN="/", MARGIN=2, STATS = colSums(MM))

#Matrix Priority (MP)
MP <- (rowSums(MN))/n 
MP <- matrix (MP, nrow=n, ncol=1)

#AX, Lambida_max, CR
AX <- (MM)%*%(MP)

Lamb_Max  <- (AX/MP)
Lamb_Max <- (sum(Lamb_Max))/n

CI = (Lamb_Max-n)/(n-1)

# RI 
 if (n==2) {RI <- 0}
  if (n==3) {RI <- 0.58}
   if (n==4) { RI <- 0.9}
   if (n==5) {RI <- 1.12}
    if (n==6) {RI <- 1.24}
      if (n==7) {RI <- 1.32}
        if (n==8) {RI <- 1.41}
          if (n==9) {RI <- 1.45}
            if (n==10) {RI <- 1.51}
              if (n >10) {RI <- NA}
 

CR=CI/RI


return (list(CR, MP))
}

```


Testing fuction

```{r testing_function_ahp}


value <- AHP(data = c(1,1,1/3,1,1,1,3,1,1), n=3)

value

```

Now think about a row with many.... 








CR - Mean, Variance and SD for each expert groupA

```{r dataset_library}

#library(readr)
#library (matrixStats)
AHP_both_CR <- read_csv("D:/Dropbox/ESTUDOS/erica phD/Field-Work-SBC/AHP-Data/AHP_both_CR.csv")

AHP_CR2 <- read_csv("D:/Dropbox/ESTUDOS/erica phD/methodology/AHP/AHP-BGM-Phases/AHP_R_IndivB.csv")

```


```{r}

# Wrangling Data:

colnames (AHP_both_CR) <- c("I", "CRA", "CRB", "CRB1", "CRB2", "CRC", "CRC1", "CRC2", "CRC3", "CRC4_", "CR3A", "CR4", "CR", "Median", "VAR.A", "VAR.P", "VAR", "SD")
AHP <- select (AHP_both_CR, CRA:CR)

#Computing Mean, Var, SD
AHP <- AHP[1:11,]
AHP <- mutate(AHP, Mean = rowMeans(AHP))
AHP <- mutate(AHP, Var = apply(AHP[,1:12], 1, var, na.rm = TRUE))
AHP <- mutate(AHP, sd = apply(AHP[,1:12], 1, sd, na.rm = TRUE))
AHP

```

CR - Mean, Variance and SD for each expert groupb

```{r}


# Wrangling Data:

colnames (AHP_CR2) <- c("I", "CRA", "CRB", "CRB1", "CRB2", "CRC1", "CRC2", "CRC3", "CRC4_", "CR3A", "CR4", "CR")
AHP_2 <- select (AHP_CR2, CRA:CR)

#Computing MEan, Var, Sd
AHP_2 <- AHP_2[1:12,]
AHP_2 <- mutate(AHP_2, Mean = rowMeans(AHP_2, na.rm	
= TRUE))
AHP_2 <- mutate(AHP_2, Var = apply(AHP_2[,1:11], 1, var, na.rm = TRUE))
AHP_2 <- mutate(AHP_2, sd = apply(AHP_2[,1:11], 1, sd, na.rm = TRUE))
AHP_2


```





1. Create Matrix
```{r matrix}

matrix_02 <- matrix (c(P1), nrow=3, ncol=3, byrow = TRUE)
matrix_priority <- matrix (Priority, nrow=3, ncol=1)


matrix_02
matrix_priority



```
**Priority**: weight for each parameter


## Checking for consistency

- Next step: compute the Consistency Ratio (CR) --> measure how consistent the judgments have been relative to the samples of purely random judgments
- AHP Assumption: decision maker is rational
- CR > 0.1 --> the judgments are untrustworthy (too close for comfort to randomness and the exercise is valuesless or must be repeated)

2. Compute Lambida Max & Average

```{r multiplication_matrix}

AX <- (matrix_02)%*%(matrix_priority)
AX

Lamb_Max  <- (AX/matrix_priority)
Lamb_Max <- (sum(Lamb_Max))/n
Lamb_Max

```
3. Compute Consistency Index (CI)

```{r CI}
n=3
CI = (Lamb_Max-n)/(n-1)
CI

```


4. Consistancy Ratio

```{r rconsistency_ratio}

RI = 0.58 ### value from table
#RI = CI/CR
CR=CI/RI
CR
```

