---
title: "AHP Brazilian Government Methodology"
author: "Erica Goto"
date: "12/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library (dplyr)
```

## AHP Brazilian Goverment Methodology
### Part 01 - Anthrop, Natural or Instability

Table


```{r dataset}

#1 - Inconsi
#A1 = c(1,1,1)
#N1 = c(1,1,1)
#I1 = c(1,1,1)

#2
A1 = c(1,0.2,0.143)
N1 = c(5,1,0.2)
I1 = c(7,5,1)






#A1 = c(1,5,7)
#N1 = c(0.2, 1, 5)
#I1 = c(0.143, 0.2, 1)



```


```{r matrix_1}





n=3

df_01 = data.frame(A1, N1, I1)

colnames(df_01) <- c("Antrop", "Natural", "Instab" )
row.names (df_01) <- c("Antrop_r", "Natural_r", "Instab_r" )
df_01

Sum <- colSums (df_01)
Sum
#row.names (df_01) <- c("Antrop_r", "Natural_r", "Instab_r")

#df_01S <- rbind(df_01, Sum)

#row.names (df_01S) <- c("Antrop_r", "Natural_r", "Instab_r", "Sum")


df_01S <- sweep(df_01, FUN="/", MARGIN=2, STATS = colSums(df_01))


df_01_N <- mutate(df_01S, Priority= rowSums(df_01S)/n) # Priority average of the row sum
df_01_N 

#Multiplication Weight for each colummn and weight sum
Priority <- df_01_N$Priority
df_01_MW <- df_01*Priority
df_01_MW
```

**Priority Column**: Weight for each parameter (row)

###  Checking for consistency

1. Create Matrix
```{r matrix}

matrix_01 <- matrix (c(A1, N1, I1), nrow=3, ncol=3)
matrix_priority <- matrix (df_01_N$Priority, nrow=3, ncol=1)


matrix_01
matrix_priority



```
**Priority**: weight for each parameter


## Checking for consistency

- Next step: compute the Consistency Ratio (CR) --> measure how consistent the judgments have been relative to the samples of purely random judgments
- AHP Assumption: decision maker is rational
- CR > 0.1 --> the judgments are untrustworthy (too close for comfort to randomness and the exercise is valuesless or must be repeated)

2. Compute Lambida Max & Average

```{r multiplication_matrix}

AX <- (matrix_01)%*%(matrix_priority)
AX

Lamb_Max  <- (AX/matrix_priority)
Lamb_Max <- (sum(Lamb_Max))/n
Lamb_Max

```
3. Compute Consistency Index (CI)

```{r CI}
n=3
CI = (Lamb_Max-n)/(n-1)
CI

```


4. Consistancy Ratio

```{r rconsistency_ratio}

RI = 0.58 ### value from table
#RI = CI/CR
CR=CI/RI
CR
```

