---
title: "AHP Brazilian Government Methodology"
author: "Erica Goto"
date: "12/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library (dplyr)
```

## AHP Brazilian Goverment Methodology
### Part 01 - Anthrop, Natural or Instability

Table


```{r dataset}

#1 - okay
#P1=c(1,1,1,1,1,1,1,1,1)
#2 
#P1=c(1,5,0.2, 0.2, 1, 0.1428,5,7,1)
#3
#P1=c(1,5,0.2,0.2,1,0.2,5,5,1)
#4
#P1=c(1,5,0.2,0.2,1,3,8,0.333333,1)

#5
#P1 = c(1,1,7,1,1,0.2, 0.14285714, 5, 1)

#6
#P1 = c(1,1,0.2,1,1,1,5,1,1)
#7
#P1 =c(1,5,0.16666667, 0.2, 1, 0.1666667,6,6,1)


#8 - okay
#P1 = c(1,1,4,1,1,4,0.25,0.25,1)

#9
#P1 = c(1,0.5,0.1111111, 0.5,1,0.2,9,5,1)

#10 - almost okay
P1 =c(1,1,0.333333, 1,1,1,3,1,1)

#11
#P1=c(1,8,8,8,1,6,0.125,0.1666667,1)


```


```{r matrix_1}





n=3

matrix_01 <- matrix(P1, nrow=3, ncol=3, byrow = TRUE)
#df_01 = data.frame(P1)

#colnames(df_01) <- c("Natural", "Antrop", "Instab" )
#row.names (df_01) <- c("Natura_rl", "Antrop_r", "Instab_r" )
matrix_01

Sum <- colSums (matrix_01)
Sum
#row.names (df_01) <- c("Antrop_r", "Natural_r", "Instab_r")

#df_01S <- rbind(df_01, Sum)

#row.names (df_01S) <- c("Antrop_r", "Natural_r", "Instab_r", "Sum")


matrix_N <- sweep(matrix_01, FUN="/", MARGIN=2, STATS = colSums(matrix_01))
matrix_N
Priority <- (rowSums(matrix_N))/n
Priority


#matrix_M <- mutate(matrix_01, Priority= rowSums(matrix_N)/n) # Priority average of the row sum
#df_01_N 

#Multiplication Weight for each colummn and weight sum
#df_01_MW <- df_01*Priority
#df_01_MW
```

**Priority Column**: Weight for each parameter (row)

###  Checking for consistency

1. Create Matrix
```{r matrix}

matrix_02 <- matrix (c(P1), nrow=3, ncol=3, byrow = TRUE)
matrix_priority <- matrix (Priority, nrow=3, ncol=1)


matrix_02
matrix_priority



```
**Priority**: weight for each parameter


## Checking for consistency

- Next step: compute the Consistency Ratio (CR) --> measure how consistent the judgments have been relative to the samples of purely random judgments
- AHP Assumption: decision maker is rational
- CR > 0.1 --> the judgments are untrustworthy (too close for comfort to randomness and the exercise is valuesless or must be repeated)

2. Compute Lambida Max & Average

```{r multiplication_matrix}

AX <- (matrix_02)%*%(matrix_priority)
AX

Lamb_Max  <- (AX/matrix_priority)
Lamb_Max <- (sum(Lamb_Max))/n
Lamb_Max

```
3. Compute Consistency Index (CI)

```{r CI}
n=3
CI = (Lamb_Max-n)/(n-1)
CI

```


4. Consistancy Ratio

```{r rconsistency_ratio}

RI = 0.58 ### value from table
#RI = CI/CR
CR=CI/RI
CR
```

